A Data-Driven Framework for Neural Field Modeling

### 2011-01-15 ltx/first_order_framework/LaTeX_Manuscript.tex
minor edits to future work

\dean{(Reviewer 2, comment 6:) The work of \citet{Freeman2000} provides direct evidence that the \parham{electrode} spatial frequency of the neural field is in the range of 0.1-0.4 cycles/mm and spacing required to prevent aliasing is approximately 1.25 mm. Micro-electrodes systems that satisfy this requirement are now currently in for brain computer interface and epilepsy monitoring applications~\citep{Cheung2007}, which suggests that the framework can be applied to humans with existing technology.}

@@

\dean{The assumption was made that the parameters of the model are stationary over the estimation period. However, it is well known that factors that influence excitability are non-stationary over longer time periods (see \citet{Ullah2010}). Therefore, the estimation framework must be altered, to account for time-varying parameters, if it is to be applied to data over extended periods of time. A possible way to achieve this is to use a more sophisticated model that accounts for longer-term time dependancies (higher temporal order). Alternatively, the parameters of the model could be considered time-varying, where changes in excitability would result in different connectivity kernel parameters. Furthermore, the model may be considered heteroscedastic, where the disturbance is also time-varying. This is common in spatio-temporal econometric studies, where methods have been developed to test for heteroscedasticity~\citep{Anselin1988}. Following this, the disturbance properties of the system will also need to be estimated. A simple method to allow for time-varying parameters would be to perform a window-based or fixed-lag estimation, which would assume the parameters are quasi-stationary over the duration of the window. }

@@

\section{Conclusion}
This paper has presented a framework for creating data-driven neural field models by developing a method for estimating intracortical connectivity and synaptic dynamics from electrophysiological data. In doing so, we have highlighted important aspects of the experiment design process, demonstrating the relationship between sensor bandwidth and the complexity of the field that can be inferred. We have derived a finite-order state-space representation of the continuum neural field equations that allows inference to be performed efficiently. Importantly, the mathematical formulation that we have presented scales well with the number of sensors. The major contribution of this framework is to formally bridge a gap between experimental and theoretical neuroscience in order to gain a better understanding of the mesoscopic dynamics of the brain. \dean{A validation of the estimation framework on real data is now required. This is the subject of planned future work.}

>>>>

\dean{(Reviewer 2, comment 6:) The work of \citet{Freeman2000} provides direct evidence that the \parham{electrode} spatial frequency of the neural field is in the range of 0.1-0.4 cycles/mm and spacing required to prevent aliasing is approximately 1.25 mm. Micro-electrodes systems that satisfy this requirement are now currently in for brain computer interface and epilepsy monitoring applications~\citep{Cheung2007}, which suggests that the framework can be applied to human data with existing technology.}

@@

\dean{The assumption was made that the parameters of the model are stationary over the estimation period. However, it is well known that factors that influence excitability are non-stationary over longer time periods (see \citet{Ullah2010}). Therefore, the estimation framework must be altered if it is to be applied to data over extended periods of time. A possible way to achieve this is to use a more sophisticated model that accounts for longer-term time dependancies (higher temporal order). Alternatively, the parameters of the model could be considered time-varying, where changes in excitability would result in different connectivity kernel parameters. Furthermore, the model may be considered heteroscedastic, where the disturbance is also time-varying. This is common in spatio-temporal econometric studies, where methods have been developed to test for heteroscedasticity~\citep{Anselin1988}. Following this, the disturbance properties of the system will also need to be estimated. A simple approach allowing for time-varying parameters is to perform a window-based or fixed-lag estimation, which would require the assumption that the parameters are quasi-stationary over the duration of the window. }

@@

\section{Conclusion}
This paper has presented a framework for creating data-driven neural field models by developing a method for estimating intracortical connectivity and synaptic dynamics from electrophysiological data. In doing so, we have highlighted important aspects of the experiment design process, demonstrating the relationship between sensor bandwidth and the complexity of the field that can be inferred. We have derived a finite-order state-space representation of the continuum neural field equations that allows inference to be performed efficiently. Importantly, the mathematical formulation that we have presented scales well with the number of sensors. The major contribution of this framework is to formally bridge a gap between experimental and theoretical neuroscience in order to gain a better understanding of the mesoscopic dynamics of the brain. \dean{A validation of the estimation framework on real data is now required.}
